---
title: "HW7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r echo=FALSE}
setwd("D:/workfiles/439/")
```

```{r echo=FALSE}
data = read.table("D:/workfiles/439/m-ibm3dx2608.txt",header = TRUE)
```

```{r echo=FALSE}
vw.ts = ts(data$vwrtn, start = 1926, frequency = 12)
```

# 1(a)
```{r echo=FALSE}
dimension_reduction = function(vw.ts,n,model_type="AR"){
  x = 0
  se_list = c()
  fix_list = rep(NA,n+1)
  current_aic = Inf
  drop_index = 0
  statistics = c()
  removal_list = c()

  while (x == 0){
    
    if (model_type == "AR"){
    AR9 = arima(vw.ts, order = c(9,0,0),fixed = fix_list)
    }else if (model_type == "MA"){
    AR9 = arima(vw.ts, order = c(0,0,9),fixed = fix_list)
    }
    
    matrix = AR9$var.coef
    dimension = dim(matrix)[1]
    for (i in 1:(dimension)){

      se_list[i] = sqrt(matrix[i,i])
    }
    
    model_aic = AR9$aic
    if (model_aic <= current_aic){
      current_aic = model_aic
      best_model = AR9
    }
    
    coeffs = coef(AR9)
    matrix = AR9$var.coef
    dimension = dim(matrix)[1]

    
    
    statistics = coeffs/se_list
    removal_list = abs(statistics[abs(statistics) < 2])
    removal_list = removal_list[removal_list !=0]
    
    for (j in 1:length(removal_list))
    if (length(removal_list) == 0){
      return(best_model)
    }
    
    drop_index = which(abs(statistics) == min(removal_list))
    removal_list = removal_list[-drop_index]
    fix_list[drop_index] = 0
  }
}
```

```{r echo=FALSE}
-2995.07/length(vw.ts)
```

```{r echo=FALSE}
arima(vw.ts, order = c(9,0,0))
```

```{r echo=FALSE}
-3000.23/length(vw.ts)
```

```{r echo=FALSE}
modelAR = dimension_reduction(vw.ts,9)
modelAR
```

```{r echo=FALSE}
-3000.23/length(vw.ts)
```

```{r echo=FALSE}
Box.test(modelAR$residuals,type = "Ljung",lag = 12,fitdf = 10)
```

## Answer: since the p-value is 0.1825 and which is greater than 0.05. The model fitted well. Compared with vw.ar9.1, their aic are same which is -3000.23. The performance of these two models are same.

# 1 (b)
```{r echo=FALSE}
modelMA = dimension_reduction(vw.ts,9,"MA")
modelMA
```

```{r echo=FALSE}
-3002.28/length(vw.ts)
```

```{r echo=FALSE}
Box.test(modelMA$residuals,type = "Ljung",lag = 12,fitdf = 10)
```
## Answer: since the p-value is 0.6302 and which is greater than 0.05. The model fitted well. Compared with vw.ar9.3, the aic of vw.ar9.3 is -3005.03 and the new model's aic is -3002.28 which is bigger than vw.ar9.3. Thus, vw.ar9.3 has a better performance. 


# 1 (c)
```{r echo=FALSE}
ARMA55 = arima(vw.ts, order = c(5,0,5))
ARMA55
```

```{r echo=FALSE}
    coeffs = coef(ARMA55)
    matrix = ARMA55$var.coef
    dimension = dim(matrix)[1]
    se_list = c()
    
    for (i in 1:(dimension)){

      se_list[i] = sqrt(matrix[i,i])
    }
    statistics = coeffs/se_list
    removal_list = abs(statistics[abs(statistics) < 2])
    removal_list
```

```{r echo=FALSE}
ARMA55 = arima(vw.ts, order = c(5,0,5),fixed = c(NA, NA, 0, NA, NA, NA, NA, NA, NA, NA, NA))
```

```{r echo=FALSE}
ARMA55
```


```{r echo=FALSE}
    coeffs = coef(ARMA55)
    matrix = ARMA55$var.coef
    dimension = dim(matrix)[1]
    se_list = c()
    
    for (i in 1:(dimension)){

      se_list[i] = sqrt(matrix[i,i])
    }
    statistics = coeffs/se_list
    removal_list = abs(statistics[abs(statistics) < 2])
    removal_list
```

```{r echo=FALSE}
ARMA55 = arima(vw.ts, order = c(5,0,5),fixed = c(NA, NA, 0, NA, NA, NA, NA, NA, 0, NA, NA))
ARMA55
```
```{r echo=FALSE}
-3008.46/length(vw.ts)
```


```{r echo=FALSE}
    coeffs = coef(ARMA55)
    matrix = ARMA55$var.coef
    dimension = dim(matrix)[1]
    se_list = c()
    
    for (i in 1:(dimension)){

      se_list[i] = sqrt(matrix[i,i])
    }
    statistics = coeffs/se_list
    removal_list = abs(statistics[abs(statistics) < 2])
    removal_list
```

```{r echo=FALSE}
ARMA55 = arima(vw.ts, order = c(5,0,5),fixed = c(NA, NA, 0, 0, NA, NA, NA, NA, 0, NA, NA))
ARMA55
```


```{r echo=FALSE}
    coeffs = coef(ARMA55)
    matrix = ARMA55$var.coef
    dimension = dim(matrix)[1]
    se_list = c()
    
    for (i in 1:(dimension)){

      se_list[i] = sqrt(matrix[i,i])
    }
    statistics = coeffs/se_list
    removal_list = abs(statistics[abs(statistics) < 2])
    removal_list
```



```{r echo=FALSE}
ARMA55 = arima(vw.ts, order = c(5,0,5),fixed = c(NA, NA, 0, 0, NA, NA, NA, 0, 0, NA, NA))
ARMA55
```


```{r echo=FALSE}
    coeffs = coef(ARMA55)
    matrix = ARMA55$var.coef
    dimension = dim(matrix)[1]
    se_list = c()
    
    for (i in 1:(dimension)){

      se_list[i] = sqrt(matrix[i,i])
    }
    statistics = coeffs/se_list
    removal_list = abs(statistics[abs(statistics) < 2])
    removal_list
```

```{r echo=FALSE}
ARMA55 = arima(vw.ts, order = c(5,0,5),fixed = c(NA, NA, 0, 0, NA, NA, NA, 0, 0, 0, NA))
ARMA55
```

```{r echo=FALSE}
    coeffs = coef(ARMA55)
    matrix = ARMA55$var.coef
    dimension = dim(matrix)[1]
    se_list = c()
    
    for (i in 1:(dimension)){

      se_list[i] = sqrt(matrix[i,i])
    }
    statistics = coeffs/se_list
    removal_list = abs(statistics[abs(statistics) < 2])
    removal_list
```

```{r echo=FALSE}
ARMA55 = arima(vw.ts, order = c(5,0,5),fixed = c(NA, NA, 0, 0, 0, NA, NA, 0, 0, 0, NA))
ARMA55
```


```{r echo=FALSE}
    coeffs = coef(ARMA55)
    matrix = ARMA55$var.coef
    dimension = dim(matrix)[1]
    se_list = c()
    
    for (i in 1:(dimension)){

      se_list[i] = sqrt(matrix[i,i])
    }
    statistics = coeffs/se_list
    removal_list = abs(statistics[abs(statistics) < 2])
    removal_list
```


```{r echo=FALSE}
ARMA55 = arima(vw.ts, order = c(5,0,5),fixed = c(0, NA, 0, 0, 0, NA, NA, 0, 0, 0, NA))
ARMA55
```


```{r echo=FALSE}
    coeffs = coef(ARMA55)
    matrix = ARMA55$var.coef
    dimension = dim(matrix)[1]
    se_list = c()
    
    for (i in 1:(dimension)){

      se_list[i] = sqrt(matrix[i,i])
    }
    statistics = coeffs/se_list
    removal_list = abs(statistics[abs(statistics) < 2])
    removal_list
```

```{r echo=FALSE}
ARMA55 = arima(vw.ts, order = c(5,0,5),fixed = c(NA, NA, 0, 0, 0, 0, NA, 0, 0, 0, 0))
ARMA55
```

```{r echo=FALSE}
Box.test(ARMA55$residuals,type = "Ljung",lag = 12,fitdf = 10)
```
